---
title: "Report"
author: "Camila and Jessica"
date: "2024-03-26"
output: html_document
---

```{r setup, include=FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))  # Setting CRAN mirror

```


```{r load_data, include=FALSE}
# Loading packages
library(ggplot2)
install.packages("tidyverse")


# Loading data
framingham_data <- read.csv("data/data_orig/framingham.csv")  

colnames(framingham_data)[colnames(framingham_data) == "prevalentHyp"] <- "pHyp"

# Removing rows with "NA" values
cleaned_data <- na.omit(framingham_data)

print(colnames(framingham_data))

```



# **INTRODUCTION**

In this analysis, we will investigate the association between body mass index (BMI) and the prevalence of hypertension among participants in the Framingham Heart Study. Obesity, often determined by BMI, is recognized as one of the principal risk factors for hypertension. Numerous studies have demonstrated that the prevalence of hypertension increases with rising BMI, therefore, understanding the relationship between BMI and hypertension prevalence is important for identifying individuals at risk, developing effective prevention as well as management strategies for hypertension. 



# **METHODS**


For this study we want to understand the association between BMI and the prevalence of hypertension on the Framingham Heart Study data set.

* ## Cleaning the Data
The Framingham data set we used was pre-processed but still included missing values (NA's) and because we do not want these NA's to affect our results we decided remove the NA's and this reduced the number of our observations. Additionally, we changed the column name of our dependent variable from 'PrevalentHYP' to 'pHyp' for clarity.


* ## Generating and Testing Hypothesis


   **Null Hypothesis**: There is no association between BMI and the prevalence of hypertension among the studied population.

   **Alternative Hypothesis**: There is an association between BMI and the prevalence of hypertension among the studied population. 
 
To test these hypotheses, we visualized the associations between BMI and hypertension using boxplots and computed the BMI median for hypertensive and non-hypertensive groups. Furthermore, we employed a logistic regression model and a t-test to assess the statistical significance of our findings.
 
 
# **RESULTS**

## Boxplot and median


```{r cleaned_data_boxplot, echo=FALSE}

# Box plot for pHyp and BMI

boxplot(BMI ~ pHyp, data = cleaned_data, 
        xlab = "Prevalent Hypertension", ylab = "BMI",
        main = "Box Plot of BMI by Prevalent Hypertension")
```


```{r cleaned_data_test, echo=FALSE}

# Filtering the dataset to include only individuals with hypertension
pHyp_data <- cleaned_data[cleaned_data$pHyp == 1, ]

# Calculate the median BMI for hypertensive individuals
median_bmi_pHyp <- median(pHyp_data$BMI)

# Print the result
print(median_bmi_pHyp)


# Filtering the dataset to include only individuals without hypertension
nopHyp_data <- cleaned_data[cleaned_data$pHyp == 0, ]

# Calculating the median BMI for individuals without hypertension
median_bmi_nopHyp <- median(nopHyp_data$BMI)

# Print the result
print(median_bmi_nopHyp)

```

The boxplot comparison shows a median BMI of 27 for the Hypertensive group and 24.67 for the Non-Hypertensive group. This suggests a potential association between higher BMI and hypertension.


## Logistic regression model and t-test


```{r cleaned_data, echo=FALSE}
#logistic regression 


logistic_model <- glm(pHyp ~ BMI, data = cleaned_data, family = binomial)

logistic_model

# data for hypertensive and non-hypertensive groups
pHyp_bmi <- cleaned_data$BMI[cleaned_data$pHyp == 1]
nopHyp_bmi <- cleaned_data$BMI[cleaned_data$pHyp == 0]

# Performing t-test
t_test_result <- t.test(pHyp_bmi, nopHyp_bmi)
print(t_test_result)
```

The logistic regression model indicates that for every 1 unit increase in BMI, Hypertension increases by 0.1707 and the t-test reveals our p-value is < 2.2e-16 which is less than 0.05 at a 95% confidence level indicating a statistically significant difference in BMI between hypertensive and non-hypertensive groups and this could not have occurred by random chance alone. Therefore, we have sufficient evidence to reject the null hypothesis in favour of the alternative hypothesis.
