---
title: "Report"
author: "Camila and Jessica"
date: "2024-03-26"
output: html_document
---

```{r setup, include=FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))  # Setting CRAN mirror

```


```{r load_data, include=FALSE}
# Load necessary packages
library(ggplot2)
install.packages("tidyverse")


# Load data
framingham_data <-read.csv("data/data_orig/framingham.csv")  

colnames(framingham_data)[colnames(framingham_data) == "prevalentHyp"] <- "pHyp"

# Remove rows with "NA" values
cleaned_data <- na.omit(framingham_data)

print(colnames(framingham_data))

```



# **INTRODUCTION**

In this analysis, we will investigate the association between body mass index (BMI) and the prevalence of hypertension among participants in the Framingham Heart Study. Obesity, often determined by BMI, is recognized as one of the principal risk factors for hypertension [1]. Numerous studies have demonstrated that the prevalence of hypertension increases with rising BMI [2,3], therefore, understanding the relationship between BMI and hypertension prevalence is important for identifying individuals at risk, developing effective prevention as well as management strategies for hypertension. 



# **METHODS**

The data from the Framingham Heart Study database was collected from kaggle.com/datasets. The data was visualiesd using a boxplot and logistic regression and t-test analysis were computed. 


* ## Cleaning the Data
The Framingham data set we used was pre-processed but still included missing values (NA's) and because we do not want these NA's to affect our results we decided remove the NA's and this reduced the number of our observations from 4240 to 3658. Additionally, we changed the column name of our dependent variable from 'PrevalentHYP' to 'pHyp' for clarity.


* ## Generating and Testing Hypotheses


   **Null Hypothesis**: There is no association between BMI and the prevalence of hypertension among the studied population.

   **Alternative Hypothesis**: There is an association between BMI and the prevalence of hypertension among the studied population. 
 
To test these hypotheses, we visualized the associations between BMI and hypertension using boxplots and computed the BMI median for hypertensive and non-hypertensive groups. Furthermore, we employed a logistic regression model and a t-test to assess the statistical significance of our findings.
 
 
# **RESULTS**

## Boxplot and median


```{r cleaned_data_boxplot, echo=FALSE}

# Box plot for pHyp and BMI

boxplot(BMI ~ pHyp, data = cleaned_data, 
        xlab = "Prevalent Hypertension", ylab = "BMI",
        main = "Box Plot of BMI by Prevalent Hypertension")
```


```{r cleaned_data_test, echo=FALSE}

# Filtering the dataset to include only individuals with hypertension
pHyp_data <- cleaned_data[cleaned_data$pHyp == 1, ]

# Calculate the median BMI for hypertensive individuals
median_bmi_pHyp <- median(pHyp_data$BMI)

# Print the result
print(median_bmi_pHyp)


# Filtering the dataset to include only individuals without hypertension
nopHyp_data <- cleaned_data[cleaned_data$pHyp == 0, ]

# Calculating the median BMI for individuals without hypertension
median_bmi_nopHyp <- median(nopHyp_data$BMI)

# Print the result
print(median_bmi_nopHyp)

```



## Logistic regression model and t-test


```{r cleaned_data, echo=FALSE}
#logistic regression 


logistic_model <- glm(pHyp ~ BMI, data = cleaned_data, family = binomial)

logistic_model

# data for hypertensive and non-hypertensive groups
pHyp_bmi <- cleaned_data$BMI[cleaned_data$pHyp == 1]
nopHyp_bmi <- cleaned_data$BMI[cleaned_data$pHyp == 0]

# Performing t-test
t_test_result <- t.test(pHyp_bmi, nopHyp_bmi)
print(t_test_result)
```



# **DISCUSSION**
The boxplot comparison shows a median BMI of 27 for the Hypertensive group and 24.67 for the Non-Hypertensive group. This suggests a potential association between higher BMI and hypertension.
From the results, the logistic regression model indicates that for every 1 unit increase in BMI, Hypertension increases by 0.1707 and the t-test reveals our p-value is < 2.2e-16 which is less than 0.05 at a 95% confidence level indicating a statistically significant difference in BMI between hypertensive and non-hypertensive groups and this could not have occurred by random chance alone. Therefore, we have sufficient evidence to reject the null hypothesis in favor of the alternative hypothesis.


# **CONCLUSION**

 Our analysis reveals a correlation between BMI values and hypertension prevalence. However, it's crucial to know that while our findings highlight an association, determining a causal effect requires further study. Factors such as confounders and unaccounted variables may influence the outcome. Additionally, adhering to the established BMI classifications (healthy weight range (18.5 - 24.9), overweight (25.0 - 29.9) and obese 30.0 or higher), emphasizes the importance of maintaining a healthy weight range for mitigating the risk of hypertension and promoting overall well-being.
 
 
# **REFERENCES**

[1] Kuwabara M., Kuwabara R., Niwa K., Hisatome I., Smits G., Roncal-Jimenez C.A., MacLean P.S., Yracheta J.M., Ohno M., Lanaspa M.A., et al. Different Risk for Hypertension, Diabetes, Dyslipidemia, and Hyperuricemia According to Level of Body Mass Index in Japanese and American Subjects. Nutrients. 2018;10:1011. doi: 10.3390/nu10081011.

[2] Lee C.Y., Lin W.T., Tsai S., Hung Y.C., Wu P.W., Yang Y.C., Chan T.F., Huang H.L., Weng Y.L., Chiu Y.W., et al. Association of Parental Overweight and Cardiometabolic Diseases and Pediatric Adiposity and Lifestyle Factors with Cardiovascular Risk Factor Clustering in Adolescents. Nutrients. 2016;8:567. doi: 10.3390/nu8090567. 

[3] Crawford A.G., Cote C., Couto J., Daskiran M., Gunnarsson C., Haas K., Haas S., Nigam S.C., Schuette R. Prevalence of obesity, type II diabetes mellitus, hyperlipidemia, and hypertension in the United States: Findings from the GE Centricity Electronic Medical Record database. Popul. Health Manag. 2010;13:151–161. doi: 10.1089/pop.2009.0039.

[4] Centers for Disease Control and Prevention - CDC. Assessing Your Weight - Health Weight, Nutrition, and Physical Activity. Accessed: 27/03/2024. Source: https://www.cdc.gov/healthyweight/assessing/index.html#print
